---
title: "R for Data Science Chapter 3: Data Transformation"
author: "Neil Martin"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true                # Enable Table of Contents
    toc_depth: 3             # Set depth of the TOC
    toc_float: true          # Move TOC
    number_sections: true    # Number the sections
    theme: cosmo             # Use a Bootstrap theme
    highlight: tango 
---

### Prerequisites

```{r setup}
library(nycflights13)
library(tidyverse)
```

### dply basics

```{r example}

glimpse(flights)

flights |>
  filter(dest == "IAH") |>
  group_by(year, month, day) |>
  summarise(
    arr_delay = mean(arr_delay, na.rm = TRUE)
  )

```

### Rows

```{r rows}

# Flights that are more than 120 minutes late

flights |>
  filter(dep_delay > 120)

# Flights that departed on January 1st

flights |>
  filter(month == 1 & day == 1)

# Flights that departed in January or February

flights |>
  filter(month == 1 | month == 2)

# A shorter way to do the above

flights |>
  filter(month %in% c(1, 2))

jan1 <- flights |>
  filter(month == 1 & day == 1)

```

### arrange()

```{r arrange}

flights |>
  arrange(year, month, day, dep_time)

flights |> 
  arrange(desc(dep_delay))

```
### distinct()

```{r distinct}

# Remove any duplicate rows

flights |>
  distinct()

# Find all unique origin and destination pairs

flights |> 
  distinct(origin, dest)

flights |>
  count(origin, dest, sort = TRUE)

```
### Exercises

#### In a single pipeline for each condition, find all flights that meet the condition:

```{r exercise1}

flights |>
  filter(arr_delay > 120,
         dest %in% c("IAH", "HOU"),
         carrier %in% c("UA", "AA", "Dl"),
         month %in% c(7,8,9),
         arr_delay > 120 & dep_delay <=0,
         dep_delay >= 60 & (dep_delay - arr_delay) >= 30)

```
#### Sort flights to find the flights with the longest departure delays. Find the flights that left earliest in the morning.

```{r exercise2}

flights |> 
  arrange(desc(arr_delay)) |>
  head(10)

flights |>
  arrange(time_hour) |>
  head(10)

```

#### Sort flights to find the fastest flights. (Hint: Try including a math calculation inside of your function.)

```{r exercise3}

flights |>
  arrange((dep_time - arr_time)) |>
  head(10)

```
#### Was there a flight on every day of 2013?

```{r exercise4}

flights |>
  mutate(date = as.Date(time_hour)) |>
  group_by(date) |>
  summarise(n_flights = n()) |>
  summarise(total_days = n()) |>
  pull(total_days)
```

<p> Yes, there was a flight everyday in 2013. </p>

#### Which flights traveled the farthest distance? Which traveled the least distance?

```{r exercsie5}

flights |>
  arrange(distance)

```

<p> Flight 51 traveled the farthest distance and flight 1632 traveled the least. </p>

#### Does it matter what order you used filter() and arrange() if you’re using both? Why/why not? Think about the results and how much work the functions would have to do.

<p> I'd argue it is better practice to filter your dataset first before applying the arrange function. This reduces the numbers of rows that will be sorted by the arrange function. Doing this later seems inefficient.</p>