---
title: "R for Data Science Chapter 7: Data Import"
author: "Neil Martin"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true                # Enable Table of Contents
    toc_depth: 3             # Set depth of the TOC
    toc_float: true          # Move TOC
    number_sections: true    # Number the sections
    theme: cosmo             # Use a Bootstrap theme
    highlight: tango 
---

### Prerequisites

```{r setup}

library(tidyverse)

```

### Reading data from a file

```{r reading_data}

students <- read_csv("https://pos.it/r4ds-students-csv")

students

students <- read_csv("https://pos.it/r4ds-students-csv", na = c("N/A", ""))

students

students |> 
  rename(
    student_id = 'Student ID',
    full_name = 'Full Name'
  )

students |> janitor::clean_names()

students

students |> 
  janitor::clean_names() |> 
  mutate(meal_plan = factor(meal_plan))

students <- students |> 
  janitor::clean_names() |> 
  mutate(
    meal_plan = factor(meal_plan),
    age = parse_number(if_else(age == "five", "5", age))
  )

students

read_csv(
  "a,b,c
  1,2,3
  4,5,6"
)

read_csv(
  "The first line of metadata
  The second line of metadata
  x,y,z
  1,2,3",
  skip = 2
)

read_csv(
  "# A comment I want to skip
x,y,z
1,2,3",
comment = "#"
)

read_csv(
  "1,2,3
  4,5,6",
  col_names = FALSE
)

read_csv(
  "1,2,3
  4,5,6",
  col_names = c("x", "y", "z")
)

```

### Exercises 

1. What function would you use to read a file where the fields were separated with "|"?

<p> I would use read_delim() to see if it automatically guesses the delimiter. </p>

2. Apart from file, skip and comment, what other arguments do read_csv() and read_tsv() have in common?

<p> The ability to be able to skip, comment and determine what col_names should be. </p>

3. What are the most important arguments to read_fwf()?

<p> The fwf_widths() and fwf_positions(). </p>

4. Sometimes strings in a CSV file contain commas. To prevent them from causing problems, they need to be surrounded by a quoting character, like " or '. By default, read_csv() assumes that the quoting character will be ". To read the following text into a data frame, what argument to read_csv() do you need to specify?

<p> quote = "'" </p>

5. Identify what is wrong with each of the following inline CSV files. What happens when you run the code?


```{r}
read_csv("a,b\n1,2,3\n4,5,6")
read_csv("a,b,c\n1,2\n1,2,3,4")
read_csv("a,b\n\"1")
read_csv("a,b\n1,2\na,b")
read_csv("a;b\n1;3")
```
6. Practice referring to non-syntactic names in the following data frame by:

Extracting the variable called 1.
Plotting a scatterplot of 1 vs. 2.
Creating a new column called 3, which is 2 divided by 1.
Renaming the columns to one, two, and three.

```{r}

annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)

extracted_1 <- annoying$'1'
extracted_1

ggplot(annoying, aes(x = '1', y = '2')) +
  geom_point() +
  labs(x = "1", y = "2", title = "Scatterplot of 1 vs 2")

annoying3 <- annoying |> 
  mutate(`3` = `2` / `1`)
annoying3

annoying_column <- annoying3 |> 
  rename(one = `1`, two = `2`, three = `3`)
annoying_column


```

### Controlling column types

```{r}
read_csv("
  logical,numeric,date,string
  TRUE,1,2021-01-15,abc
  false,4.5,2021-02-15,def
  T,Inf,2021-02-16,ghi
")

simple_csv <- "
x
10
.
20
30"

read_csv(simple_csv)

df <- read_csv(
  simple_csv,
  col_types = list(x= col_double())
)

problems(df)

read_csv(simple_csv, na = ".")

another_csv <- "
x,y,z
1,2,3"

read_csv(
  another_csv,
  col_types = cols(.default = col_character())
)

read_csv(
  another_csv,
  col_types = cols_only(x = col_character())
)

```

### Reading data from multiple files

```{r}



```




