---
title: "R for Data Science Chapter 7: Data Import"
author: "Neil Martin"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true                # Enable Table of Contents
    toc_depth: 3             # Set depth of the TOC
    toc_float: true          # Move TOC
    number_sections: true    # Number the sections
    theme: cosmo             # Use a Bootstrap theme
    highlight: tango 
---

### Prerequisites

```{r setup}

library(tidyverse)

```

### Reading data from a file

```{r reading_data}

students <- read_csv("https://pos.it/r4ds-students-csv")

students

students <- read_csv("https://pos.it/r4ds-students-csv", na = c("N/A", ""))

students

students |> 
  rename(
    student_id = 'Student ID',
    full_name = 'Full Name'
  )

students |> janitor::clean_names()

students

students |> 
  janitor::clean_names() |> 
  mutate(meal_plan = factor(meal_plan))

students <- students |> 
  janitor::clean_names() |> 
  mutate(
    meal_plan = factor(meal_plan),
    age = parse_number(if_else(age == "five", "5", age))
  )

students

read_csv(
  "a,b,c
  1,2,3
  4,5,6"
)

read_csv(
  "The first line of metadata
  The second line of metadata
  x,y,z
  1,2,3",
  skip = 2
)

read_csv(
  "# A comment I want to skip
x,y,z
1,2,3",
comment = "#"
)

read_csv(
  "1,2,3
  4,5,6",
  col_names = FALSE
)

read_csv(
  "1,2,3
  4,5,6",
  col_names = c("x", "y", "z")
)

```

### Exercises 

1. What function would you use to read a file where the fields were separated with "|"?

<p> I would use read_delim() to see if it automatically guesses the delimiter. </p>

2. Apart from file, skip and comment, what other arguments do read_csv() and read_tsv() have in common?

<p> The ability to be able to skip, comment and determine what col_names should be. </p>

3. What are the most important arguments to read_fwf()?

<p> The fwf_widths() and fwf_positions(). </p>

