---
title: "R for Data Science Chapter 2: Workflow Basics"
author: "Neil Martin"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true                # Enable Table of Contents
    toc_depth: 3             # Set depth of the TOC
    toc_float: true          # Move TOC
    number_sections: true    # Number the sections
    theme: cosmo             # Use a Bootstrap theme
    highlight: tango 
---

### Loading libraries

```{r setup}

library(tidyverse)
library(ggthemes)

```

### Exercises

#### Why does this code not work?

```{r exercise1}

my_variable <- 10

my_variable

```

<p> This code didn't work due to not running the code which defines the variable beforehand.</p>

#### Tweak each of the following R commands so that they run correctly:

```{r exercise2}

library(tidyverse)

ggplot(data = mpg, aes(x = displ, y = hwy)) + 
  geom_point() +
  geom_smooth(method = "lm")


```
#### Press Option + Shift + K / Alt + Shift + K. What happens? How can you get to the same place using the menus?

<p> Help > Keyboard Shortcuts Help </p>

#### Let’s revisit an exercise from the Section 1.6. Run the following lines of code. Which of the two plots is saved as mpg-plot.png? Why?

```{r exercise4}

my_bar_plot <- ggplot(mpg, aes(x = class)) +
  geom_bar()
my_scatter_plot <- ggplot(mpg, aes(x = cty, y = hwy)) +
  geom_point()
ggsave(filename = "mpg-plot.png", plot = my_bar_plot)

```
#### Let’s revisit an exercise from the Section 1.6. Run the following lines of code. Which of the two plots is saved as mpg-plot.png? Why?

<p> The bar plot gets saved as. This is due to the ggsave function defining the plot to be the my_bar_plot variable which stores the bar plot.</p>

