---
title: "R for Data Science Chapter 1: Data Visualisation"
author: "Neil Martin"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true                # Enable Table of Contents
    toc_depth: 3             # Set depth of the TOC
    toc_float: true          # Move TOC
    number_sections: true    # Number the sections
    theme: cosmo             # Use a Bootstrap theme
    highlight: tango 
---

### Loading libraries

```{r setup}
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
```

### Exploration

```{r eda}
penguins

glimpse(penguins)
```
### Ggplot

```{r ggplot}

ggplot(data = penguins,
       mapping = aes(x= flipper_length_mm, y = body_mass_g)) +
  geom_point()

```

### Visuals

```{r vis_dist}

ggplot(penguins, aes(x = species)) +
  geom_bar()

ggplot(penguins, aes(x = fct_infreq(species))) +
  geom_bar()

```

### Numerical Variables

```{r num_var}

ggplot(penguins, aes(x = body_mass_g)) +
  geom_histogram(binwidth = 200)

ggplot(penguins, aes(x = body_mass_g)) +
  geom_histogram(binwidth = 20)

ggplot(penguins, aes(x = body_mass_g)) +
  geom_histogram(binwidth = 2000)

ggplot(penguins, aes(x = body_mass_g)) +
  geom_density()

```

### Exercises

#### Make a bar plot of species of penguins, where you assign species to the y aesthetic. How is this plot different?

```{r exercise1}

ggplot(penguins, aes(y = species)) +
  geom_bar()

```
<p> The plot is different as it is horizontal in nature compared to putting the Species on the x axis. It shows the same results but just in a different fashion.</p>

#### How are the following two plots different? Which aesthetic, color or fill, is more useful for changing the color of bars?

```{r exercise2}

ggplot(penguins, aes(x = species)) +
  geom_bar(color = "red")

ggplot(penguins, aes(x = species)) +
  geom_bar(fill = "red")

```

<p>The first bar uses the color function to color the perimeter of the bar chart while the fill function fills the bar with a specified color. The fill option would be more useful as this can be used later to outline the differences between variables in colour.</p>

### Numerical and categorical variables

```{r num_cat_var}

ggplot(penguins, aes(x = species, y = body_mass_g)) +
  geom_boxplot() +
  labs(title = "Body mass by Species (Boxplot)")

ggplot(penguins, aes(x = body_mass_g, color = species)) +
  geom_density(linewidth = 0.75) +
  labs(title = "Body mass by Species (Density)")

```

